---
title: "Migration2013"
author: "Bo Suzow"
date: "February 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The purpose of this report to explore the bilateral migration data available at [KNOMAD](https://www.knomad.org/data/migration/emigration). The Global Knowledge Partnership on Migration and Development (KNOMAD) is a global hub of knowledge and policy expertise on migration and development issues.

The dataset utilized in this report can be accessed  [here.](https://www.knomad.org/sites/default/files/2017-03/bilateralmigrationmatrix20130.xlsx)

The report extracts data pertinent to the emigration from the USA for the exploratory analysis and plotting.  [A companion Shiny app](http://www.tinyurl.com/bsuzow/MIG2013) allows the user to select a country for emigration or immigration data. 

***
## Data load and preparation

```{r, dataprep}

library(readxl)  #http://readxl.tidyverse.org/
library(dplyr)
library(data.table)
library(ggplot2)
library(plotly)


download.file("https://www.knomad.org/sites/default/files/2017-03/bilateralmigrationmatrix20130.xlsx","Mig2013.xlsx", mode="wb")
Mig2013 = read_excel("Mig2013.xlsx",range="A2:HJ219")
names(Mig2013)[1]="Country"  # shorten the col name

# Select the US for the focus of the exploratory analysis, 
#    

Mig2013US = Mig2013 %>% filter(Country == "United States")  # extracting data points pertinent to US

Mig2013US = melt(Mig2013US)
Mig2013US$Country = NULL  # remove the first col whose value is "United States" for all rows, thus meaningless

names(Mig2013US) = c("Country","NumMig")
Mig2013US = Mig2013US %>% filter(NumMig>0)

```

Now, the dataset needs to be augmented with country codes. With a quick Google search, I found [a source for Country Codes](https://raw.githubusercontent.com/plotly/datasets/master/2014_world_gdp_with_codes.csv). Unfortunately, it did not cover all countries represented in the dataset downloaded from KNOMAD. Further research led me to the countrycode R package which turned out to be a better choice.

```{r}

CountryCodes=read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2014_world_gdp_with_codes.csv")

names(CountryCodes)[1]="Country"

Mig2013US2= merge(x=Mig2013US,y=CountryCodes,by="Country",all.x=TRUE)
# ref: https://stackoverflow.com/questions/1299871/how-to-join-merge-data-frames-inner-outer-left-right
# This returned about 20 countries with no code.

# Fortunately,I found the countrycode package. 

library(countrycode)
Mig2013US2 = Mig2013US %>% mutate(Code=countrycode(Country,"country.name","iso3c"))

# A much better result.  It is still coded north korea incorrectly.  No country codes for Other North and Other South.  This should be noted in the Shiny app.

Mig2013US3 = Mig2013US2 %>% 
               filter(!is.na(Code)) %>% # remove countries with no code
               mutate(Code=ifelse(grepl("^Korea,( +)Dem",Country),"PRK",Code)) %>% # fix north korea's code
               mutate(hoverpop = paste(as.character(round(NumMig/1000,1)),
                                       "K moved to ",
                                       Country,sep="")) # compose the tooltip text


```

## Plotting - Emigration

Let's plot a world map visualizing how many US citizens have moved to other countries.

```{r, plot}


borders = list(color= toRGB("black"),width=1)
map_options = list(
   scope="world",
   projection = list(type='Mercator'),
   showcoastlines = TRUE
   ) # borders
     
p3 = plot_ly(z   = ~Mig2013US3$NumMig,
            text = ~Mig2013US3$hoverpop,
            locations = ~Mig2013US3$Code,
            type ="choropleth",
            locationmode='world',
            color=~Mig2013US3$NumMig,
            colors="Blues",
            marker = list(line=borders)) %>%
   
     colorbar(title = "Migration Outflow") %>%
   
     plotly::layout(title='US citizens migrate out to (Hover for breakdown)', geo=map_options)
     
p3

```


Sort the data frame on the migration outflow numbers in descending order to list top 10 countries the US citizens moved to.


```{r}

Mig2013US3.desc = Mig2013US3 %>% arrange(desc(NumMig))
head(Mig2013US3.desc,10)

```

***
## Companion Shiny App

As mentioned earlier in the report, you may take a look at [the Shiny app](http://www.tinyurl.com/bsuzow/MIG2013) that allows you to select a country for emigration or immgration data.




